### 2016-10-24 dissertation/prelim.tex
dissertation: milestone 1: first draft

\chapter{Computer Algebra Fundamentals} \label{ch:ideals}
This chapter reviews fundamental concepts of commutative and 
computer algebra which are used in this work. 
Specifically, this chapter covers monomial ordering, polynomial ideals and 
varieties, and the computation of \Grobner bases.
It also overviews elimination theory as well as Hilbert's Nullstellensatz 
theorems and how
they apply to Galois fields. The results of these theorems are used in polynomial
abstraction and formal verification of Galois field circuits and 
are discussed 
in subsequent chapters. The material of this chapter is
mostly referred from the textbooks \cite{ideals:book} \cite{gb_book} and 
previous work by {\it Lv} \cite{lv:phd}.

\section{Monomials, Polynomials, and Term Orderings}

\begin{Definition} \label{def:mono}
A {\bf monomial} in variables $x_1,x_2,\cdots,x_d$ is a product of the form:
\begin{equation}
x_1^{{\alpha}_1} \cdot x_2^{{\alpha}_2} \cdot \cdots x_d^{{\alpha}_d},
\end{equation}
where $\alpha_i \ge 0, i\in\{1,\cdots,d\}$. 
The total degree of the monomial is $\alpha_{1}+\cdots+\alpha_{d}$.
\end{Definition}

Thus, $x^2\cdot y$ is a monomial in variables $x,y$ with total degree $3$.
For simplicity, we will henceforth denote a 
monomial $x_1^{{\alpha}_1} \cdot x_2^
{{\alpha}_2} \cdot \cdots x_d^{{\alpha}_d}$ as $x^{\alpha}$, 
where $\alpha=({\alpha}_1,\cdots,{\alpha}_d)$ is a vector size $d$ of 
integers $\ge 0$, i.e., $\alpha \in \mathbb{Z}_{\ge 0}^{d}$. In the

@@

Since a polynomial is a sum of its terms,
these terms have to be arranged unambiguously so that they can be 
manipulated in a consistent manner.
Therefore, we need to establish a concept of 
{\bf term ordering} (also called monomial ordering).
A term ordering, represented by $>$, defines how
terms in a polynomial are ordered.
%%%%%%%%%%%%%%%%%%%   monomial    Ordering   %%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{Definition}
Let $\mathbb{T}^{d}=\{x^{\alpha}: \alpha\in \mathbb{Z}_{\ge 0}^{d}\}$ be the set of all monomials in $x_{1},\dots,x_{d}$.
A {\bf monomial order} $>$ on $\mathbb{T}^{d}$ is a total well-ordering satisfying:
\begin{itemize}
        \item For all $x^{\alpha}, x^{\beta} \in \mathbb{T}^{d}$, $x^{\alpha}$ and $x^{\beta}$ are comparable
	\item For any $x^{\alpha} \in \mathbb{T}^{d}$, $x^{\alpha}>1$
	\item For all $x^{\alpha}, x^{\beta}, x^{\gamma} \in \mathbb{T}^{d}$, $x^{\alpha}>x^{\beta} \Rightarrow x^{\alpha} \cdot x^{\gamma}> x^{\beta} \cdot x^{\gamma}$
\end{itemize}
\end{Definition}

Term-orderings are totally ordered, i.e. anti-symmetric with constant terms
last in the ordering.
A total-order ensures that there is no ambiguity with respect to where a 
term is found in the term-ordering.  Total orderings for monomials come in 
different forms, notably 
lexicographic orderings (lex), and its variants: degree-lexicographic ordering (deglex) 
and reverse degree-lexicographic ordering (degrevlex).

@@

Applying these term orderings, we have the following relations, 
where $a > b > c$.

\begin{eqnarray}
    \eqntext{lex:}  
    a^2b > a^2 > abc > ab > ac^2 > ac > b^2c > b^2 > bc^3 > 1 
    \label{ex:ordering:lex}\\
    \eqntext{deglex:} 
    bc^3 > a^2b > abc > ac^2 > b^2c > a^2 > ab >
    ac > b^2 > 1
    \label{ex:ordering:deglex}\\
    \eqntext{degrevlex:}  
    bc^3 > a^2b > abc > b^2c > ac^2 > a^2 > ab >
    b^2 > ac > 1
    \label{ex:ordering:degrevlex}
\end{eqnarray}

The difference between the {\it lex} and two {\it deg-} orderings is
obvious, while the difference between the two degree-based orderings can be
seen by considering from which direction the term is lexed, e.g. $a\cdot c\cdot c > b\cdot b\cdot c$ (deglex, left-to-right)
versus $b\cdot b\cdot c > a \cdot c\cdot c$ (degrevlex, right-to-left).

\begin{Example}
Let $f = 2x^2yz + 3xy^3 - 2x^3$. The effects of different term orderings on 
$f$ are:
\begin{itemize}
\item lex $x> y> z$: $f = -2x^3 + 2x^2yz + 3xy^3$
\item deglex $x>y>z$:  $f = 2x^2yz + 3xy^3 -2x^3$
\item degrevlex $x>y>z$: $f = 3xy^3 + 2x^2yz - 2x^3$
\end{itemize}
\end{Example}

%Based on the {\it monomial ordering}, we have the following concepts:

@@

\section{Varieties and Ideals}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%  variety %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In computer-algebra based formal verification,
it is often necessary to analyze the presence 
or absence of solutions to a given system of constraints.
In our applications, these constraints are polynomials and their solutions 
are modeled as {\bf varieties}.

@@

When $J = \sqrt J$, then $J$ is said to be a 
{\it radical ideal}. Moreover, $I(V)$ is a radical ideal.
By analyzing the ideal $J$, generated by a system of polynomials derived 
from a hardware design, its variety $V(J)$, and the ideal of 
polynomials that vanish over this variety, $I(V(J))$, we can reason about the 
existence of certain properties of the design. To check for the existence of 
a property, we formulate the 
property as a polynomial and then perform an {\bf ideal membership test} to 
determine if this polynomial is contained within the ideal $I(V(J))$. 
A {\bf \Grobner basis} provides a decision procedure for performing this 
test, which is described in the following section. 
A future section focuses on {\bf Hilbert's Nullstellensatz}, 
which describes the properties of the ideal of a variety, $I(V(J))$.

%%%%%%%%%%%%%%%%%%%%	Grobner bases	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\Grobner Bases}

@@

\begin{Definition}
    The {\bf reduction} of a polynomial $f$, by another polynomial $g$, to
    a reduced polynomial $r$ is denoted:
    \begin{equation*}
        f\stackrel{g}{\textstyle\longrightarrow}r
    \end{equation*}
    Reduction is carried out using multivariate, polynomial long division. 
    % The long division is performed according to a term-ordering on polynomials, and the division algorithm 
    %terminates when the leading term of the divisor does not divide any 
    %other term in the dividend.

@@

\section{Elimination Theory}

@@

Elimination theory provides the basis for our abstraction approach.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Hilbert's Nullstellensatz}

In this section, we further describe some correspondence between ideals and 
varieties in the context of algebraic geometry. The celebrated results of 
Hilbert's Nullstellensatz establish these correspondences.

%%%%%%%%%%%%%%%%%algebraically closed field%%%%%%%%%%%%%
\begin{Definition}\label{def:acf}
A field $\overline {\K}$ is an {\bf algebraically closed} field if every  
polynomial in one variable with degree at least $1$, with coefficients 
in $\overline {\K}$, has a root in $\overline {\K}$. 
\end{Definition}
In other words, any non-constant polynomial equation over 
$\overline {\K}\left[x\right]$ always has at least one root 
in $\overline {\K}$. Every field $\K$ is contained in an algebraically 
closed one $\overline {\K}$. 
For example, the field of real numbers $\mathbb{R}$ is not an algebraically closed 
field, because $x^2+1=0$ has no root in $\mathbb{R}$. 
However, $x^2+1=0$ has roots in the field of 
complex numbers $\mathbb{C}$, which is an algebraically closed field. 
In fact, $\mathbb{C}$ is the algebra closure of $\mathbb{R}$. 
Every algebraically closed field is an infinite field.

>>>>

\chapter{Preliminaries}
\label{ch:prelim}
This chapter provides a mathematical background for understanding 
Galois fields and explains how to design Galois field arithmetic circuits.
Specifically, this chapter covers monomial ordering, polynomial ideals and 
varieties, and the computation of \Grobner bases.
Moreover, we introduce the mathematical concepts of groups, rings, fields, and 
polynomials. 
We then apply these concepts to create Galois field arithmetic functions and 
explain how elimination and abstraction on Galois field works.
The material is referred from \cite{galois_field:mceliece} \cite{ftheory:2006} \cite{ff:1997} for Galois field concepts,
\cite{ideals:book} \cite{gb_book} for commutative algebra basics and 
previous work by {\it Lv} \cite{lv:phd} and {\it Pruss} \cite{tim:phd}.

\section{Commutative Algebra}
\label{sec:algebra}
\subsection{Group, ring and field}
\begin{Definition}
An {\bf abelian group} is a set $\mathbb{S}$ with a binary operation $'+'$
which satisfies the following properties: 
\begin{itemize}
\item {\it Closure Law:} For every $a, b \in \mathbb{S}, a + b \in \mathbb{S}$  
\item {\it Associative Law:} For every $a, b, c \in \mathbb{S}, (a + b) + c = a + (b + c)$
\item {\it Commutativity:} For every $a, b \in \mathbb{S}, a + b = b + a$. 
\item {\it Additive Identity:} There is an identity element $0 \in \mathbb{S}$
such that for all $a \in \mathbb{S};$ $a + 0 = a$.
\item {\it Additive Inverse:} If $a \in \mathbb{S}$, then there is an
element $a^{-1} \in \mathbb{S}$ such that $ a + a^{-1} = 0$.
\end{itemize}
\end{Definition}

The set of integers $\mathbb{Z}$ forms an abelian group under the addition operation.

\begin{Definition}
Given a set $\mathbb{R}$ with two binary operations, $'+'$ and $'\cdot'$, 
and element $0 \in \mathbb{R}$, the system $\mathbb{R}$ is called a {\bf commutative ring with unity} if the following properties hold:
\begin{itemize}
\item $\mathbb{R}$ forms an abelian group under the '+' operation with additive identity element $0$.
\item {\it Multiplicative Distributive Law}: For all $a, b, c \in$ $\mathbb{R}$, $a\cdot (b + c) = a\cdot b + a\cdot c$.
\item {\it Multiplicative Associative Law}: For every $a, b, c\in \mathbb{R}$, $a\cdot (b\cdot c) = (a\cdot b)\cdot c$. 
\item {\it Multiplicative Commutative Law}: For every $a,b \in \mathbb{R}$, $a\cdot b = b\cdot a$
\item {\it Identity Element}: There exists an element $1 \in$ $\mathbb{R}$ 
such that for all $a \in \mathbb{R}$, $a\cdot 1 = a =1\cdot a$
\end{itemize}
\end{Definition}

For the purpose of this dissertation, any time we refer to a {\bf ring}, we are 
specifically referring to a {\bf commutative ring with unity}. Two common 
examples of such rings are the set of integers, $\mathbb{Z}$, and the set of 
rational numbers, $\mathbb{Q}$. Note that while both of these examples are
rings with an infinite number of elements, the number of elements in a ring 
can also be finite.

\begin{Definition}
A {\bf field} $\mathbb{F}$ is a commutative ring with unity, where every
non-zero element in $\mathbb{F}$ has a multiplicative inverse; i.e. $\forall
a \in \mathbb{F} - \{0\}$, $\exists \hat{a} \in \mathbb{F}$ such that $ a \cdot
\hat{a} = 1$.
\end{Definition}

A field is defined as a ring with one extra condition: the presence of a 
multiplicative inverse for all non-zero elements.
Therefore, a field must be a ring while a ring is not necessarily a field.
For example, the set $\mathbb{Z}_{2^k} = \{0,1,\cdots, 2^k-1\}$ forms a finite ring.
However, $\mathbb{Z}_{2^k}$ is not a field because not every element in
$\mathbb{Z}_{2^k}$ has a multiplicative inverse. 
In the ring $\mathbb{Z}_{2^3}$, for 
instance, the element $5$ has an inverse ($5\cdot5\pmod{8}=1$) but the element $4$
does not.

The main concept of field theory is {\bf Field Extensions}. The idea behind a
field extension is to take a base field and construct a larger field which 
contains the base field as well as satisfies additional properties. For example,
the set of real numbers $\mathbb{R}$ forms a field; one common extension of 
$\mathbb{R}$ is the set of complex numbers $\mathbb{C}=\mathbb{R}(i)$. Every
element of $\mathbb{C}$ can be represented as $a+b\cdot i$ where $a,b \in \mathbb{R}$,
hence $\mathbb{C}$ is a two-dimensional extension of $\mathbb{R}$.

Like rings, fields can also contain either an infinite or a finite number of 
elements. 
In this dissertation we focus on finite fields, also known as Galois fields, and 
the construction of their field extensions.

\subsection{Galois Field}
Galois fields, also known as finite fields, find widespread applications in 
many areas of electrical engineering and computer science such as error-
correcting codes, elliptic curve cryptography, digital signal processing, 
testing of VLSI circuits, among others.
In this dissertation, we specifically focus on their application to 
Elliptic Curve Cryptography as Galois field arithmetic circuits.
This section describes the relevant Galois field concepts
\cite{galois_field:mceliece} \cite{ftheory:2006} \cite{ff:1997}
and hardware arithmetic designs over such fields \cite{mastro:1989} \cite{PT:1985} 
\cite{acar:1998} \cite{wu:2002} \cite{Knezevic:2008}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{Definition} 
A {\bf Galois field}, denote $\Fq$, is a field with a finite
number of elements, $q$. The number of elements $q$ of the Galois field is
a power of a prime integer, i.e. $q = p^k$, where $p$ is a prime
integer, and $k \geq 1$. Thus a Galois field can also be denoted as 
$\F_{p^{k}}$.
\end{Definition}

Fields in the form $\F_{p^{k}}$ are called Galois extension fields.
We are specifically interested in extension fields of type 
$\Fkk$, where $k > 1$. These are extensions of the binary
field $\F_2$.
\begin{Example}
Addition and multiplication operations over $\F_2$:
\begin{table}[!h]
	\centering
	\begin{tabular}{m{1cm}|l|ll|m{1cm}}
	\hhline{~---~}
	\multirow{3}{*}{} & $+$ & $0$ & $1$ & \multirow{3}{*}{} \\
	\hhline{~---~}
	& $0$ & $0$ & $1$ & \\
	& $1$ & $1$ & $0$ & \\
	\hhline{~---~}
	\multicolumn{5}{c}{}\\
	\multicolumn{5}{c}{Addition over $\F_2$}\\
	\end{tabular}
	\quad
	\begin{tabular}{m{1cm}|l|ll|m{1cm}}
	\hhline{~---~}
	\multirow{3}{*}{} & $\cdot$ & $0$ & $1$ & \multirow{3}{*}{} \\
	\hhline{~---~}
	& $0$ & $0$ & $0$ & \\
	& $1$ & $0$ & $1$ & \\
	\hhline{~---~}
	\multicolumn{5}{c}{}\\
	\multicolumn{5}{c}{Multiplication over $\F_2$}\\
	\end{tabular}
\end{table}

Notice that addition over $\F_2$ is a Boolean {\sc XOR} operation, 
because it is performed modulo $2$.
Similarly, multiplication over $\F_2$ performs a Boolean {\sc AND} operation.
\end{Example}

Algebraic extensions of the binary field $\F_{2}$  
are generally termed as {\it binary extension fields} $\Fkk$.
Where elements in $\F_2$ can only represent $1$ bit, elements in $\Fkk$ 
represent a $k$-bit vector.
This allows them to be widely used in digital hardware applications.
In order to construct a Galois field of the form $\Fkk$, 
an {\bf irreducible polynomial} is required:
\begin{Definition}
A polynomial $P(x) \in \mathbb{F}_{2}\left[x\right]$ is {\bf irreducible} 
if $P(x)$ is non-constant with degree $k$ and cannot be 
factored into a product of polynomials of lower degree in $\mathbb{F}_2[x]$.
\end{Definition}

Therefore, the polynomial $P(x)$ with degree $k$ is irreducible over 
$\mathbb{F}_{2}$ if and only if it has no roots in $\mathbb{F}_{2}$,
i.e if $\forall a \in \mathbb{F}_{2}$, $P(a)\neq 0$.
For example, $x^2+x+1$ is an irreducible polynomial over $\mathbb{F}_{2}$
because it has no solutions in $\mathbb{F}_{2}$, i.e. $(0)^2+(0)+1=1\neq0$ 
and $(1)^2+(1)+1=1\neq0$ over $\F_2$.
Irreducible polynomials exist for any degree $\geq 2$ in $\mathbb{F}_2[x]$.

Given an irreducible polynomial $P(x)$ of degree $k$ in the polynomial ring 
$\mathbb{F}_2[x]$, we can construct a binary extension field 
$\mathbb{F}_{2^k} \equiv \mathbb{F}_2[x] \pmod{P(x)}$.
Let $\alpha$ be a root of $P(x)$, i.e., $P(\alpha)=0$.
Since $P(x)$ is irreducible over
$\mathbb{F}_2[x]$, $\alpha \notin \mathbb{F}_2$. 
Instead, $\alpha$ is an element in $\mathbb{F}_{2^k}$. 
Any element $A \in \mathbb{F}_{2^k}$ is then represented as: 
\begin{equation}\label{rep:poly}
A= \sum_{i=0}^{k-1} (a_i \cdot \alpha^i) = a_0 + a_1\cdot\alpha + \cdots + a_{k-1}\cdot \alpha^{k-1}\nonumber
\end{equation}
where $a_i \in \mathbb{F}_2$ are the coefficients and $P(\alpha)=0$.

To better understand this field extension, compare its similarities to another
common-place
field extension $\C$, the set of complex numbers. $\C$ is an extension of the field 
of real numbers $\R$ with an additional element $i=\sqrt{-1}$, which is an imaginary
root in $\R$.
Thus $i \notin \R$, rather $i \in \C$.
Every element $A \in \mathbb{C}$ can be represented as:
\begin{equation}\label{rep:polyC}
A=\sum_{j=0}^{1} (a_j \cdot i^j)=a_0+a_1\cdot i
\end{equation}
where $a_j \in \R$ are coefficients. Similarly, $\Fkk$ is an extension of $\F_2$ with 
an additional element $\alpha$, which is the ``imaginary root'' of an irreducible 
polynomial $P$ in $\F_2[x]$.

Every element $A \in \Fkk$ has a degree less than $k$ because 
$A$ is always computed modulo $P(x)$, which has degree $k$. 
Thus, $A\pmod {P(x)}$ can be of degree at most $k-1$ and at least $0$.
For this reason, the field $\mathbb{F}_{2^k}$ can be viewed as a $k$
dimensional vector space over $\mathbb{F}_{2}$. 
The equivalent bit vector representation for element $A$ is:
\begin{equation}
A=(a_{k-1} a_{k-2} \cdots a_{0})
\end{equation}

\begin{Example}
A 4-bit Boolean vector, $(a_{3} a_{2} a_{1} a_{0})$
can be presented over $\F_{2^4}$ as: 
\begin{equation}
a_3 \cdot \alpha^3+a_2 \cdot \alpha^2+a_1 \cdot \alpha+a_0
\end{equation}
For instance, the Boolean vector $1011$ is represented as the element 
$\alpha^3+\alpha+1$.
\end{Example}

\begin{Example}\label{exp:1}
Let us construct $\mathbb{F}_{2^4}$ as $\mathbb{F}_2[x] \pmod{ P(x)}$, where
$P(x)=x^4+x^3+1 \in \mathbb{F}_2[x]$ is an irreducible polynomial of degree $k=4$. 
Let $\alpha$ be the root of $P(x)$, i.e. $P(\alpha)=0$.

Any element $A \in \mathbb{F}_2[x] \pmod{ x^4 + x^3 + 1}$
has a representation of the type: $A = a_3 x^3 + a_2 x^2 +
a_1 x + a_0$ (degree $< 4$) where the coefficients $a_3, \dots, a_0$ are in $\F_2 =
\{0, 1\}$. Since there are only $16$ such polynomials, we obtain
$16$ elements in the field $\mathbb{F}_{2^4}$. Each element in
$\mathbb{F}_{2^4}$ can then be viewed as a $4$-bit vector over $\mathbb{F}_{2}$. 
Each element also has an exponential $\alpha$
representation. All three representations are shown in Table
\ref{tab:gfelement}.

\begin{table}[h]
\begin{center}
\caption{Bit-vector, Exponential and Polynomial representation of
elements in  $\mathbb{F}_{2^4} = \mathbb{F}_2[x] \pmod{x^4+x^3+1}$}\label{tab:gfelement} 
\begin{tabular}{|c|c|c||c|c|c|} 
\hline
$a_3a_2a_1a_0$ & Exponential & Polynomial     &$a_3a_2a_1a_0$ & Exponential & Polynomial  \\
\hline
$0000$        & $0$         & $0$            & $1000$ & $\alpha^3$ &  $\alpha^3$\\
\hline
$0001$        & $1$         & $1$            & $1001$ & $\alpha^4$ & $\alpha^3 + 1$\\
\hline
$0010$        & $\alpha$    & $\alpha$       & $1010$ & $\alpha^{10}$&$\alpha^3 + \alpha$  \\
\hline
$0011$        & $\alpha^{12}$& $\alpha + 1$   & $1011$ & $\alpha^5$ & $\alpha^3+\alpha+1$\\
\hline
$0100$        & $\alpha^2$  & $\alpha^2$     &  $1100$ & $\alpha^{14}$ & $\alpha^3 + \alpha^2$\\
\hline
$0101$        & $\alpha^9$   &$\alpha^2 + 1$ & $1101$  &$\alpha^{11}$  & $\alpha^3+\alpha^2+1$\\
\hline
$0110$        & $\alpha^{13}$& $\alpha^2 + \alpha$ & $1110$ & $\alpha^8$& $\alpha^3+\alpha^2+\alpha$\\
\hline
$0111$        &$\alpha^7 $ & $\alpha^2+\alpha+1$ & $1111$ &$\alpha^6$ & $\alpha^3+\alpha^2+\alpha+1$\\
\hline
\end{tabular}
\end{center}
\end{table}

We can compute the polynomial representation from the exponential representation.
Since every element is computed $\pmod{P(\alpha)} = \pmod{\alpha^4+\alpha^3+1}$, 
we compute the element $\alpha^{4}$ as 
\begin{equation}
\alpha^{4} \pmod{ \alpha^4+\alpha^3+1} = -\alpha^3 - 1 = \alpha^3+1
\end{equation}
Recall that all coefficients of $\F_{2^4}$ 
are in $\F_{2}$ where $-1 = +1$ modulo 2.
The next element $\alpha^{5}$ can be computed as 
\begin{equation}
\alpha^{5} = \alpha^{4}\cdot \alpha = (\alpha^3+1)\cdot \alpha = \alpha^4+\alpha = \alpha^3+\alpha+1 
\end{equation}
Then $\alpha^6$ can be computed as $\alpha^{5}*\alpha$ and so on.
\end{Example}

An irreducible polynomial can also be a primitive polynomial.

\begin{Definition}
A {\bf primitive polynomial} $P(x)$ is a polynomial with coefficients in $\mathbb{F}_2$ 
which has a root $\alpha$ $\in$ $\mathbb{F}_{2^k}$
such that \{$0$, $1(=\alpha^{{2^k}-1})$, $\alpha$, $\alpha^2$, $\cdots$, $\alpha^{2^k-2}$\} is the set of 
all elements in $\mathbb{F}_{2^k}$, 
where $\alpha$ is a {\bf primitive element} of $\mathbb{F}_{2^k}$. 
\end{Definition}

A primitive polynomial is guaranteed to generate all distinct elements 
of a finite field $\mathbb{F}_{2^k}$ while an irreducible polynomial
has no such guarantee.
Often, there exists more than one irreducible polynomial of degree $k$.
In such cases, any degree $k$ irreducible polynomial can be 
used for field construction. For example, both $x^3+x+1$ and $x^3+x^2+1$ 
are irreducible in $\mathbb{F}_2$ and either one can be used
to construct $\mathbb{F}_{2^3}$. This is due to the following:

\begin{Theorem}\label{the:unique}
There exist a {\bf unique} field $\mathbb{F}_{p^k}$, for any prime $p$ and any positive integer $k$.
\end{Theorem}

Theorem \ref{the:unique} implies that Galois fields with the same number of elements are 
{\bf isomorphic} to each other up to the labeling of the elements.

Theorem \ref{the:fer} provides an important property for investigating solutions to
polynomial equations in $\Fq$.

\begin{Theorem}\label{the:fer}
 $\left[Generalized\  Fermat's\  Little\  Theorem \right]$ Given a
 Galois field $\mathbb{F}_{q}$, each element $A \in \mathbb{F}_{q}$ satisfies: 
\begin{eqnarray}\label{fe}
 A^{q} & \equiv & A  \nonumber \\
 A^{q} - A & \equiv& 0  
\end{eqnarray}
\end{Theorem}

We can extend Theorem \ref{the:fer} to polynomials in $\mathbb{F}_{q}[x]$ as 
follows: 
\begin{Definition}
Let $x^q-x$ be a polynomial in $\mathbb{F}_{q}[x]$.
Every element $A \in \mathbb{F}_{q}$ is a solution to  $x^q-x=0$. 
Therefore, $x^{q} - x$ always {\it vanishes} in $\mathbb{F}_{q}$. Such 
polynomials are called {\bf vanishing polynomials} of the field $\mathbb{F}_{q}$.
\end{Definition}

\begin{Example}
Given $\mathbb{F}_{2^2} =\{0,1,\alpha,\alpha+1\}$ with $P(x)=x^2+x+1$, where $P(\alpha)=0$. 
 \begin{eqnarray}
 0^{2^2}&=&0 \nonumber \\
 1^{2^2}&=&1 \nonumber \\
 \alpha^{2^2}&=&\alpha \pmod {\alpha^2+\alpha+1}\nonumber \\
 (\alpha+1)^{2^2}&=&\alpha+1 \pmod {\alpha^2+\alpha+1} \nonumber 
 \end{eqnarray}
\end{Example}

A Galois field $\F_q$ can be fully contained within a larger field $\F_{q^k}$.
That is, $\F_q \subset \F_{q^k}$.
For example, Fig \ref{fig:contain2_4_16} shows the containment of the fields 
$\F_2 \subset \F_4 \subset \F_{16}$. It's easy to see that since $\F_4=\F_{2^2}$, it
contains $\F_2$. Likewise $\F_{16}=\F_{4^2}=\F_{2^4}$ contains $\F_4$ and $\F_2$.
The elements $\{0,1,\alpha,\dots,\alpha^{14}\}$
designate $\F_{16}$. Of these, $\{0,1,\alpha^5,\alpha^{10}\}$ create $\F_4$.
From these, only $\{0,1\}$ exist in $\F_2$.

\begin{Theorem}
$\F_{2^n}\subset\F_{2^m}$ iff $n \mid m$, i.e. if $n$ divides $m$.
\end{Theorem}

Therefore:
\begin{itemize}
\item $\F_2 \subset \F_{2^2} \subset \F_{2^4} \subset \F_{2^8} \subset \dots$
\item $\F_2 \subset \F_{2^3} \subset \F_{2^9} \subset \F_{2^{27}} \subset \dots$
\item $\F_2 \subset \F_{2^5} \subset \F_{2^{25}} \subset \F_{2^{125}} \subset \dots,$ and so on
\end{itemize}

\begin{Definition}
The {\bf algebraic closure} of the Galois field $\F_{2^k}$, denoted $\overline{\Fkk}$, is the 
union of all fields $\F_{2^n}$ such that $k \mid n$.
\end{Definition}

% TODO \Fkk operations example

\subsection{Monomials, Polynomials and Polynomial Arithmetics}
\begin{Definition} \label{def:mono}
A {\bf monomial} in variables $x_1,x_2,\cdots,x_d$ is a product of the form:
\begin{equation}
x_1^{{\alpha}_1} \cdot x_2^{{\alpha}_2} \cdot \cdots x_d^{{\alpha}_d},
\end{equation}
where $\alpha_i \ge 0, i\in\{1,\cdots,d\}$. 
The total degree of the monomial is $\alpha_{1}+\cdots+\alpha_{d}$.
\end{Definition}

Thus, $x^2\cdot y$ is a monomial in variables $x,y$ with total degree $3$.
For simplicity, we will henceforth denote a 
monomial $x_1^{{\alpha}_1} \cdot x_2^
{{\alpha}_2} \cdot \cdots x_d^{{\alpha}_d}$ as $x^{\alpha}$, 
where $\alpha=({\alpha}_1,\cdots,{\alpha}_d)$ is a vector size $d$ of 
integers $\ge 0$, i.e., $\alpha \in \mathbb{Z}_{\ge 0}^{d}$.

\begin{Definition}\label{def:poly}
Let $\mathbb{R}$ be a ring. A {\bf polynomial} over $\mathbb{R}$ in the 
indeterminate $x$ is an expression of the form:
\begin{equation} \label{eq:poly1}
a_0 + a_1 x + a_2 x^2 + \cdots + a_k x^k = \sum_{i=0}^{k} a_i x^i, \forall a_i \in \mathbb{R}. 
\end{equation}

The constants $a_i$ are the coefficients and $k$ is the degree of the polynomial. 
For example, $8x^3 + 6x + 1$ is a polynomial in $x$ over $\mathbb{Z}$, with
coefficients $8$, $6$, and $1$ and degree $3$.

\begin{Definition}
The set of all polynomials in the indeterminate
$x$ with coefficients in the ring $\mathbb{R}$ forms a {\bf
ring of polynomials} $\mathbb{R}[x]$. 
Similarly, $\mathbb{R}[x_1,x_{2},\cdots, x_{n}]$ 
represents the ring of multivariate polynomials with coefficients in $\mathbb{R}$.
\end{Definition}

For example, $\mathbb{Z}_{2^4}[x]$ stands for the set of all polynomials in
$x$ with coefficients in $\mathbb{Z}_{2^4}$. $8x^3 + 6x + 1$ is an instance 
of a polynomial contained in $\mathbb{Z}_{2^4}[x]$.

@@

Since a polynomial is a sum of its terms,
these terms have to be arranged unambiguously so that they can be 
manipulated in a consistent manner.
Therefore, we need to establish a concept of 
{\bf term ordering} (also called monomial ordering).
A term ordering, represented by $>$, defines how
terms in a polynomial are ordered.

Common term orderings are lexicographic ordering (LEX) and its variants: 
degree-lexicographic ordering (DEGLEX) and reverse degree-lexicographic ordering (DEGREVLEX).

@@

Based on the {\it monomial ordering}, we have the following concepts:

@@

Similarly, we can also define when a polynomial is divided (reduced) by a set of polynomials.
\begin{Definition}
    The {\bf reduction} of a polynomial $f$, by another polynomial $g$, to
    a reduced polynomial $r$ is denoted:
    \begin{equation*}
        f\stackrel{g}{\textstyle\longrightarrow}r
    \end{equation*}
    Reduction is carried out using multivariate, polynomial long division. 
    % The long division is performed according to a term-ordering on polynomials, and the division algorithm 
    %terminates when the leading term of the divisor does not divide any 
    %other term in the dividend.

\section{Algebraic Geometry}
\label{sec:geo}
\subsection{Varieties and Ideals}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%  variety %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In computer-algebra based formal verification,
it is often necessary to analyze the presence 
or absence of solutions to a given system of constraints.
In our applications, these constraints are polynomials and their solutions 
are modeled as {\bf varieties}.

@@

When $J = \sqrt J$, then $J$ is said to be a 
{\it radical ideal}. Moreover, $I(V)$ is a radical ideal.
By analyzing the ideal $J$, generated by a system of polynomials derived 
from a hardware design, its variety $V(J)$, and the ideal of 
polynomials that vanish over this variety, $I(V(J))$, we can reason about the 
existence of certain properties of the design. To check for the existence of 
a property, we formulate the 
property as a polynomial and then perform an {\bf ideal membership test} to 
determine if this polynomial is contained within the ideal $I(V(J))$. 
A {\bf \Grobner basis} provides a decision procedure for performing this 
test, which is described in the following part.

\subsection{\Grobner Bases}

@@

@@

\section{Elimination and Abstraction Theory}
\label{sec:abstraction}
\subsection{Elimination Theory}

@@

Elimination theory provides the basis for following abstraction approach.
\subsection{Abstraction using Nullstellensatz and Gr\"obner Basis}
Let $S$ be the system of 
polynomials, 
$\{f_1,\dots,f_s,f_{A_1},\dots,f_{A_n},f_{Z}\}\subset \Fkk$, 
derived from the hardware
implementation of the Galois field arithmetic circuit over $\Fkk$.
This circuit performs some unknown function $f$ over 
$\Fkk$ in the form of $Z=\Func(A_1,\dots,A_n)$, where $Z$ is the $k$-bit 
output and $A_1,\dots,A_n$ are the $k$-bit inputs.
The polynomial representation of $\Func$ over $\Fkk$ is thus:

\begin{equation}
f_{\Func}: Z+\Func(A_1,\dots,A_n) \nonumber
\end{equation}

Since $f_{\Func}$ is ultimately derived from the circuit implementation, 
it agrees with the solution to the system of polynomials $\{S\}=0$, i.e.:
\begin{equation}
f_1=\dots=f_s=f_{A_1}=\dots=f_{A_n}=f_{Z}=0 \nonumber
\end{equation}
Thus, if we let $J=\langle f_1,\dots,f_s,f_{A_1},\dots,f_{A_n},f_{Z}\rangle$ 
be the ideal generated by $S$, 
$f_{\Func}$ {\bf vanishes} on the variety $V_{\Fkk}(J)$. 
Therefore, due to Proposition \ref{pro:iofv}, 
$f_{\Func}$ must be contained in the ideal of polynomials that vanish on
this variety, $f_{\Func} \in I(V_{\Fkk}(J))$.

By applying Strong Nullstellensatz over $\Fkk$ (Theorem \ref{thm:sns}), 
$I(V_{\Fkk}(J))=J+J_0$ where
$J_0$ is the ideal generated by all vanishing polynomials in $\Fkk$. 
Recall that a vanishing polynomial in $\Fkk[x]$ is $x^q-x=x^q+x$. 
In our case, 
$\{x_1,\dots,x_d\} \in \F_2$ and $\{A_1,\dots,A_n,Z\} \in \Fkk$.
Thus, for $\Fkk[x_1,\dots,x_d,A_1,\dots,A_n,Z]$: 
\begin{equation}
J_0 = \langle x_1^2+x_1,\dots,x_d^2+x_d, A_1^{2^k}+A_1,\dots,A_n^{2^k}+A_n,Z^{2^k}+Z\rangle \nonumber
\end{equation}

The generators of the ideal sum $J+J_0$ are simply the combination of the 
generators of $J$ and the generators $J_0$.
